---
title: "Analysis of Poorly Performing Nursing Homes within the United States"
author: "Dave Holtschlag"
date: " `r format(Sys.Date(), '%A %B %d, %Y') ` "
output: 
  html_document:
    toc: true
    toc_float: true
bibliography: bib/bibliography.bib  
---

```{r setup, include=FALSE}
library(tidyverse)
# library(kable)
library(kableExtra)
library(ggthemes)
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract
An analysis of data on poorly performing nursing homes in the United States...[@falcone_u.s._2015]

## Introduction


## Purpose
The purpose of this article is to describe the relation between attributes of provides of nursing home care compiled by Medicare and measures of the quality of that care.  Quality of care include  

### Sources of Data

#### Source of Nursing Home Provider Data
https://data.medicare.gov/Nursing-Home-Compare/Provider-Info/4pq5-n9py/data
This provider data set was exported as a TSV for Excel file and downloaded from the above site on July 18, 2019. The data set contains 15,527 rows and 85 columns of variables.


#### Source Classifying Under Performing Nursing Homes as Special Focus Facility (SFF) or SFF Candidate
https://www.aging.senate.gov/press-releases/casey-and-toomey-secure-secret-document-from-administration-list-includes-hundreds-of-poorly-performing-nursing-homes-across-the-country
This rating data set was accessed as a PDF file as link labelled 'Read the list here' (https://www.aging.senate.gov/download/special-focus-facility-list-with-cover-letter).
The dataset contains 478 rows with unique Federal Provider Numbers and 3 columns of variables

Supplimentary information was obtained from the link labelled 'Read the report from Sens. Casey and Toomey here.' (https://www.aging.senate.gov/download/casey/toomey-sff-report), the link labelled 'Read the letter to CMS from Sens. Casey and Toomey here. (https://www.aging.senate.gov/imo/media/doc/2019.3.4%20Aging%20Casey%20Toomey%20Letter%20to%20CMS%20Administrator%20re.%20Special%20Focus%20Facilities%20PA%20Final.pdf), and the link labelled 'Read the CMS response here. (https://www.aging.senate.gov/imo/media/doc/CMS%20Response%20to%20Ranking%20Member%20Casey%20050319.pdf).


#### Source of data on penalties on all US Nursing Homes
https://data.medicare.gov/Nursing-Home-Compare/All-US-Nursing-Home-Penalties/2shu-hjj2
Data was exported on July 19, 2019, in TSV for Excel format.

### Ownership 
https://data.medicare.gov/data/nursing-home-compare, accessed July 21, 2019
A list of ownership information for currently active nursing homes.

### Star Ratings based on Provider Info
https://data.medicare.gov/Nursing-Home-Compare/Star-Ratings/ax9d-vq6k, July 21, 2019
Star Ratings Based on Provider Info Overall, Health Inspection, Staffing, and Quality Measure Ratings for all active providers.


```{r read_data}

# Read in nursing home provider info
# Note: read.csv was more robust reading the table than read.table
provider <- read.csv(file = 'Data/Nursing Home Provider_Info.tsv', sep='\t',
                       header = TRUE, stringsAsFactors = FALSE,
                     colClasses = c(Provider.Zip.Code = 'character'))

# Confirm that all Provider.Numbers in the dataframe are unique
if (length(unique(provider$Federal.Provider.Number) == nrow(provider))){
  print(paste('All', nrow(provider), 'Provider.Numbers in the Provider dataset are unique.'))
} else{
  print(paste('There are',nrow(provider) - length(unique(provider$Federal.Provider.Number)),
              'rows in the Provider dateset'))
}

# Add field to get breakdown by: "For profit" "Non profit" "Government"
provider$Ownership.Type.General <- substr(provider$Ownership.Type, 1, 10)

# Read in Nursery home SFF ratings
sff_rate <- read.csv(file = 'Data/Nursing Home Rating.tsv', sep = '\t',
                     header = TRUE, stringsAsFactors = FALSE,
                     colClasses = c('character', 'character', 'factor'))


# Confirm that all Provider.Numbers in the sff_rate dataframe are unique
if (length(unique(sff_rate$Federal.Provider.Number) == nrow(sff_rate))){
  print(paste('All', nrow(sff_rate), 'Provider.Numbers in the sff_rate dataset are unique.'))
} else{
  print(paste('There are',nrow(sff_rate) - length(unique(sff_rate$Federal.Provider.Number)),
              'rows in the Provider dateset'))
}

# Left join of provider and sff_rate dataframes
sff_provider <- left_join(sff_rate, provider, by = 'Federal.Provider.Number')
print( paste('Left join of sff_rate and provider dataframes has', nrow(sff_provider), 'rows.') )

# Right join of provider and sff_rate dataframes
provider_sff <- right_join(sff_rate, provider, by = 'Federal.Provider.Number')
print( paste('Right join of sff_rate and provider dataframes has', nrow(provider_sff), 'rows.') )

# Initialize variable to contain SFF or SFF Candidate variable
provider_sff$SFF_or_Candidate <- 'None'

# Find indices of provider_sff dataframe designated SFF or SFF candidate
ndx <- which(substr(provider_sff$Special.Focus.Status, 1, 3) == 'SFF')

# Replace 'None' with SFF Status based on index
provider_sff$SFF_or_Candidate[ndx] <- 'SFF or Candidate'

# Create frequency table of SFF by State
Special.Focus.State.table <- table(provider_sff$Provider.State,
                                   provider_sff$SFF_or_Candidate)

# Convert frequency table of SFF by State to dataframe
Special.Focus.State.df    <- as.data.frame.matrix(Special.Focus.State.table)

Special.Focus.State.df$SFFc_per <- Special.Focus.State.df$`SFF or Candidate` /
  (Special.Focus.State.df$`SFF or Candidate` + Special.Focus.State.df$None) * 100

#
penalties <- read.csv(file = 'Data/All_US_Nursing_Home_Penalties.tsv',
                      sep='\t', header = TRUE, stringsAsFactors = FALSE,
                      colClasses = c('character', 'character', 'character',
                                     'character','numeric'))
penalties$Penalty.Date <- as.Date(penalties$Penalty.Date, "%m/%d/%Y")

tmp <- left_join(penalties, provider[,1:2], by = 'Provider.Name')


```

### Variation in the number of nursing homes and residents among states 

You can also embed plots, for example:

```{r freq_tables_by_state}

# Set up provider dataframe based on
provider %>% 
  group_by(Provider.State) %>% 
  summarise(no.nursing.homes.in.state = n(), 
            pop.in.state.nursing.homes = sum(Average.Number.of.Residents.Per.Day)) %>% 
  ggplot(aes(x = reorder(Provider.State, -no.nursing.homes.in.state), 
             y = no.nursing.homes.in.state)) +
  geom_bar(stat = 'identity') +
  theme_few()     +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 9)) +
  scale_y_continuous(breaks = c(2, 10, 50, 100, 200, 500, 1000, 1500), 
                     limits = c(0, 1500), trans = 'sqrt',
                     name = 'Number of Nursing Homes Tracked in State' ) +
  scale_x_discrete(name = 'State Postal Abbreviation')
  

# Summary data for population with missing values
summary_data <- provider %>% 
  group_by(Provider.State) %>% 
  summarise(no.nursing.homes.in.state = n(), 
            pop.in.state.nursing.homes = sum(Average.Number.of.Residents.Per.Day, 
                                             na.rm = TRUE),
            no.missing.nursing.homes = sum(is.na(Average.Number.of.Residents.Per.Day)))

kable(summary_data, 
      caption = 'Table 1. Number of Nursing Homes and Population Served by State') %>%
  kable_styling(bootstrap_options = "striped", fixed_thead = TRUE)  %>%
  scroll_box(width = "95%", height = "400px")

summary_data %>% 
  ggplot(aes(x = reorder(Provider.State, -pop.in.state.nursing.homes), 
             y = pop.in.state.nursing.homes)) +
  geom_bar(stat = 'identity') +
  theme_few() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 9)) +
  scale_y_continuous(
    # limits = c(1, 120000),
    breaks = c(0, 5000, 15000, 30000, 60000, 90000), 
                     trans = 'sqrt',
                     name = 'Number of Nursing Home Residents Tracked in State' ) +
  scale_x_discrete(name = 'State Postal Abbreviation')


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## References
